---
title: "results_app2"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
# If needed, set the working directory here:
# knitr::opts_knit$set(root.dir = "...")
```

# Data reading and functions

```{r}
setwd("~/Library/CloudStorage/OneDrive-UniversitàdegliStudidiMilano/OneDrive/TESI STUDENTI/TESI SEGUITE polimi/TESI BERGONZOLI 2023/J Clas/Applicazione OMERF education/ESS11MD_e01_2")
source("OMERF_ranger.R")
source("OMERF_clm_ranger.R")
# If these files are in the working directory, they will be loaded.
# If not present, the script uses fallback functions defined below.
source("ord_class_index.R")
source("index.R")

library(caret)
library(ordinal)
library(ordinalForest)
library(pdfCluster)
library(psych)
library(ranger)

#set.seed(1999)

dati <- readRDS("clean_data")

# Factors
factor_vars <- c("mainact", "gndr", "cntry")
factor_vars <- intersect(factor_vars, names(dati))
dati[factor_vars] <- lapply(dati[factor_vars], as.factor)

# Ordered response variable and group
# Note: the original variable in the dataset is stflife or happy.
dati$y <- factor(dati$stflife, ordered = TRUE)
dati$group <- factor(dati$cntry)

# Better to use these as numeric from a performance standpoint
#dati$health <- factor(dati$health, ordered=T)
#dati$ppltrst <- factor(dati$ppltrst, ordered=T)
#dati$trstplt <- factor(dati$trstplt, ordered=T)
#dati$rlgdgr <- factor(dati$rlgdgr, ordered=T)

# Covariates used in all models
xnam <- c("agea", "health", "netustm", "hinctnta", "eisced", "mainact", "gndr",
          "ppltrst", "pplfair", "pplhlp", "trstplt", "trstlgl", "rlgdgr")

# Keep only the necessary columns and remove missing values
vars_keep <- c("y", "group", xnam)
dati_model <- na.omit(dati[, vars_keep])

str(dati_model)
summary(dati_model)
table(dati_model$y)
table(dati_model$group)
length(levels(dati_model$group))

barplot(table(dati$y))
```

# Train/test split

```{r}
set.seed(1999)
split_indices <- createDataPartition(dati_model$y, p = 0.80, list = FALSE)

data_train <- dati_model[split_indices, ]
data_test  <- dati_model[-split_indices, ]

# Align factor levels between train and test
for (v in names(data_train)) {
  if (is.factor(data_train[[v]])) {
    data_test[[v]] <- factor(data_test[[v]], levels = levels(data_train[[v]]),
                             ordered = is.ordered(data_train[[v]]))
  }
}

y <- data_train$y
group <- data_train$group
cov <- data_train[, xnam]

y_test_orig <- data_test$y
y_test <- as.numeric(y_test_orig)
lev <- levels(y)
K <- length(lev)

# Check: response with ordered levels and 28 groups, as expected.
K
length(levels(group))
```

# Support functions for prediction and metrics

```{r}
# Converts a probability matrix into a predicted class.
prob_to_class <- function(prob, lev) {
  prob <- as.data.frame(prob)
  names(prob) <- lev
  factor(lev[apply(prob, 1, which.max)], levels = lev, ordered = TRUE)
}

# Manual prediction for CLMM with random intercept included.
# For groups not observed in training, b_i = 0 is used.
predict_clmm_re <- function(model, newdata, lev) {
  fixed_formula <- y ~ agea + health + netustm + hinctnta + eisced + mainact + gndr +
                       ppltrst + pplfair + pplhlp + trstplt + trstlgl + rlgdgr
  X <- model.matrix(fixed_formula, data = newdata)
  beta <- model$beta
  X <- X[, names(beta), drop = FALSE]
  eta_fixed <- as.numeric(X %*% beta)

  re <- ranef(model)$group
  b <- rep(0, dim(newdata)[1])
  for (i in 1:dim(newdata)[1]) {
    b[i] <- re[rownames(re) == as.character(newdata$group[i]), ]
  }
  b[is.na(b)] <- 0

  eta <- eta_fixed + as.numeric(b)
  theta <- model$Theta
  K <- length(lev)

  cumprob <- matrix(NA_real_, nrow = nrow(newdata), ncol = K - 1)
  for (k in 1:(K - 1)) {
    cumprob[, k] <- plogis(theta[k] - eta)
  }

  prob <- matrix(NA_real_, nrow = nrow(newdata), ncol = K)
  prob[, 1] <- cumprob[, 1]
  if (K > 2) {
    for (k in 2:(K - 1)) {
      prob[, k] <- cumprob[, k] - cumprob[, k - 1]
    }
  }
  prob[, K] <- 1 - cumprob[, K - 1]

  prob[prob < 0] <- 0
  prob <- prob / rowSums(prob)
  colnames(prob) <- lev

  list(
    mu = as.data.frame(prob),
    response = prob_to_class(prob, lev),
    eta = eta
  )
}

# CLM prediction: probabilities and predicted class.
predict_clm_prob <- function(model, newdata, lev) {

  fixed_formula <- y ~ agea + health + netustm + hinctnta + eisced + mainact + gndr +
                       ppltrst + pplfair + pplhlp + trstplt + trstlgl + rlgdgr

  X <- model.matrix(fixed_formula, data = newdata)
  beta <- model$beta

  X <- X[, names(beta), drop = FALSE]
  eta <- as.numeric(X %*% beta)

  theta <- model$Theta
  K <- length(lev)

  cumprob <- matrix(NA_real_, nrow = nrow(newdata), ncol = K - 1)

  for (k in 1:(K - 1)) {
    cumprob[, k] <- plogis(theta[k] - eta)
  }

  prob <- matrix(NA_real_, nrow = nrow(newdata), ncol = K)
  prob[, 1] <- cumprob[, 1]

  for (k in 2:(K - 1)) {
    prob[, k] <- cumprob[, k] - cumprob[, k - 1]
  }

  prob[, K] <- 1 - cumprob[, K - 1]

  prob[prob < 0] <- 0
  prob <- prob / rowSums(prob)
  colnames(prob) <- lev

  list(
    mu = as.data.frame(prob),
    response = factor(
      lev[apply(prob, 1, which.max)],
      levels = lev,
      ordered = TRUE
    )
  )
}

# Ordinal forest prediction.
predict_ordforest_prob <- function(model, newdata, lev) {
  pr <- predict(model, newdata = newdata)
  mu <- as.data.frame(pr$classprobs)
  names(mu) <- lev
  list(
    mu = mu,
    response = factor(as.character(pr$ypred), levels = lev, ordered = TRUE)
  )
}

# OMERF prediction: uses the custom method, with random effects included.
predict_omerf_prob <- function(model, y_train, newdata, group_test, lev) {
  mu <- predict.omerf(model, y_train, newdata, group_test, type = "mu")
  mu <- as.data.frame(mu)
  names(mu) <- lev
  response <- predict.omerf(model, y_train, newdata, group_test, type = "response")
  response <- factor(as.character(response), levels = lev, ordered = TRUE)
  list(mu = mu, response = response)
}

# Square confusion matrix with all levels.
confusion_full <- function(y_true, y_pred, lev) {
  tab <- table(
    factor(y_true, levels = lev, ordered = TRUE),
    factor(y_pred, levels = lev, ordered = TRUE)
  )
  dimnames(tab) <- list(observed = lev, predicted = lev)
  tab
}

# Fallback for ordinal index, if OrdinalClassificationIndex is not available.
ordinal_index_safe <- function(cm) {
  if (exists("OrdinalClassificationIndex")) {
    return(OrdinalClassificationIndex(cm, nrow(cm)))
  }
  NA_real_
}

# Fallback for newindex, if the function is not available.
newindex_safe <- function(mu, y_num, K) {
  if (exists("newindex")) {
    return(newindex(mu, y_num, K)$norm.index)
  }
  NA_real_
}

# Cohen's Weighted Kappa for ordinal response.
# weights = "quadratic" penalizes distant errors more heavily;
# use weights = "linear" for linear penalization.
weighted_kappa_safe <- function(y_true, y_pred, lev, weights = "quadratic") {
  y_true <- factor(y_true, levels = lev, ordered = TRUE)
  y_pred <- factor(y_pred, levels = lev, ordered = TRUE)

  tab <- table(y_true, y_pred)
  n <- sum(tab)
  K <- length(lev)

  if (n == 0 || K < 2) {
    return(NA_real_)
  }

  d <- abs(row(tab) - col(tab))
  if (weights == "linear") {
    W <- d / (K - 1)
  } else if (weights == "quadratic") {
    W <- (d / (K - 1))^2
  } else {
    stop("weights must be 'linear' or 'quadratic'")
  }

  O <- tab / n
  E <- outer(rowSums(tab), colSums(tab)) / n^2

  den <- sum(W * E)
  if (den == 0) {
    return(NA_real_)
  }

  1 - sum(W * O) / den
}

# Macro-averaged precision (one-vs-rest per class, then averaged)
macro_precision <- function(y_true, y_pred, lev) {
  prec <- sapply(lev, function(c) {
    tp <- sum(y_true == c & y_pred == c)
    fp <- sum(y_true != c & y_pred == c)
    if ((tp + fp) == 0) return(NA) else return(tp / (tp + fp))
  })
  mean(prec, na.rm = TRUE)
}

# Macro-averaged recall (one-vs-rest per class, then averaged)
macro_recall <- function(y_true, y_pred, lev) {
  rec <- sapply(lev, function(c) {
    tp <- sum(y_true == c & y_pred == c)
    fn <- sum(y_true == c & y_pred != c)
    if ((tp + fn) == 0) return(NA) else return(tp / (tp + fn))
  })
  mean(rec, na.rm = TRUE)
}

model_metrics <- function(name, pred, y_true, lev) {
  y_num <- as.numeric(y_true)
  pred_num <- as.numeric(pred$response)
  cm <- confusion_full(y_true, pred$response, lev)

  data.frame(
    model = name,
    accuracy = mean(y_num == pred_num),
    recall = macro_recall(y_num, pred_num, as.numeric(lev)),
    precision = macro_precision(y_num, pred_num, as.numeric(lev)),
    rmse = sqrt(mean((y_num - pred_num)^2)),
    mae = mean(abs(y_num - pred_num)),
    ordinal_classification_index = ordinal_index_safe(cm),
    adjusted_rand_index = adj.rand.index(y_num, pred_num),
    cohen_kappa = cohen.kappa(cbind(y_num, pred_num))$kappa,
    cohen_weighted_kappa = weighted_kappa_safe(
      y_true = y_true,
      y_pred = pred$response,
      lev = lev,
      weights = "quadratic"
    ),
    new_index = newindex_safe(pred$mu, y_num, length(lev))
  )
}
```

# CLM

The `clm` model uses fixed effects only, so it does not include group effects.

```{r}
clm.data <- data_train

clm1 <- clm(
  y ~ agea + health + netustm + hinctnta + eisced + mainact + gndr +
      ppltrst + pplfair + pplhlp + trstplt + trstlgl + rlgdgr,
  data = clm.data,
  link = "logit"
)
summary(clm1)

pred_clm <- predict_clm_prob(clm1, data_test, lev)
head(pred_clm$mu)
table(pred_clm$response)
```

# CLMM

The `clmm` model includes the random intercept for `group`. Prediction is computed manually to include the estimated random effects.

```{r}
clmm1 <- clmm(
  y ~ agea + health + netustm + hinctnta + eisced + mainact + gndr +
      ppltrst + pplfair + pplhlp + trstplt + trstlgl + rlgdgr + (1 | group),
  link = "logit",
  data = data_train,
  Hess = TRUE
  #, control = clmm.control(maxLineIter = 200, maxIter = 500, grtol = 1e-4)
)
summary(clmm1)
VarCorr(clmm1)

pred_clmm <- predict_clmm_re(clmm1, data_test, lev)
head(pred_clmm$mu)
table(pred_clmm$response)
```

# Ordinal forest

The same fixed covariates as in the formula are used. Here `group` is not included as a random effect.

```{r}
ordfor.data <- data_train[, c("y", xnam)]
ordforest1 <- ordfor(
  depvar = "y",
  perffunction = "probability",
  data = ordfor.data
)

pred_ordforest <- predict_ordforest_prob(ordforest1, data_test[, xnam], lev)
head(pred_ordforest$mu)
table(pred_ordforest$response)
```

# OMERF

OMERF uses the fixed covariates in the random forest and the random intercept of `group` in the mixed model.

```{r}
omerf1 <- omerf(
  y = y,
  cov = cov,
  group = group,
  xnam = xnam,
  znam = NULL,
  bizero = NULL,
  itmax = 200,
  toll = 0.02
)
summary.omerf(omerf1)
VarCorr(omerf1$clmm.model)

pred_omerf <- predict_omerf_prob(
  model = omerf1,
  y_train = y,
  newdata = data_test[, c(xnam, "group")],
  group_test = data_test$group,
  lev = lev
)
head(pred_omerf$mu)
table(pred_omerf$response)
```

# OMERF_CLM

OMERF_CLM uses the fixed covariates in the random forest and the random intercept of `group` in the mixed model.

```{r}
omerf1_clm <- omerf_clm(
  y = y,
  cov = cov,
  group = group,
  xnam = xnam,
  znam = NULL,
  bizero = NULL,
  itmax = 200,
  toll = 0.02
)
summary.omerf(omerf1_clm)
VarCorr(omerf1_clm$clmm.model)

pred_omerf_clm <- predict_omerf_prob(
  model = omerf1_clm,
  y_train = y,
  newdata = data_test[, c(xnam, "group")],
  group_test = data_test$group,
  lev = lev
)
head(pred_omerf_clm$mu)
table(pred_omerf_clm$response)
```

# Random effects

## OMERF

```{r}
ranef.omerf(omerf1, group)
```

## CLMM

```{r}
re_clmm <- ranef(clmm1)$group[, 1]

# THERE IS NO DIRECT WAY TO COMPUTE CONFIDENCE INTERVALS
#lev_group_clmm <- names(re_clmm)
#ci <- confint(clmm1, parm = "group", level = 0.95)
#ci <- ci[lev_group_clmm, , drop = FALSE]

ord.re <- order(re_clmm)

plot(
  seq_along(re_clmm),
  re_clmm[ord.re],
  axes = FALSE,
  xlab = "group",
  ylab = "group effect"
)

axis(1)
axis(2)
box()

abline(h = 0, lty = 2)
```

# Performance measures

```{r}
results <- rbind(
  model_metrics("clm",       pred_clm,       y_test_orig, lev),
  model_metrics("clmm",      pred_clmm,      y_test_orig, lev),
  model_metrics("ordforest", pred_ordforest, y_test_orig, lev),
  model_metrics("omerf",     pred_omerf,     y_test_orig, lev),
  model_metrics("omerf_clm", pred_omerf_clm, y_test_orig, lev)
)

results

results.acc  <- setNames(results$accuracy,                     results$model)
results.prec <- setNames(results$precision,                    results$model)
results.rec  <- setNames(results$recall,                       results$model)
results.rmse <- setNames(results$rmse,                         results$model)
results.mae  <- setNames(results$mae,                          results$model)
results.oc   <- setNames(results$ordinal_classification_index, results$model)
results.ari  <- setNames(results$adjusted_rand_index,          results$model)
results.ck   <- setNames(results$cohen_kappa,                  results$model)
results.wk   <- setNames(results$cohen_weighted_kappa,         results$model)
results.newi <- setNames(results$new_index,                    results$model)

results.acc  # the higher the better
results.prec # the higher the better
results.rec  # the higher the better
results.rmse # the lower the better
results.mae  # the lower the better
results.oc   # the closer to 0 the better
results.ari  # the higher the better
results.ck   # the higher the better
results.wk   # the higher the better
results.newi # the closer to 0 the better
```

# Confusion matrix

```{r}
cm_clm       <- confusion_full(y_test_orig, pred_clm$response,      lev)
cm_clmm      <- confusion_full(y_test_orig, pred_clmm$response,     lev)
cm_ordforest <- confusion_full(y_test_orig, pred_ordforest$response, lev)
cm_omerf     <- confusion_full(y_test_orig, pred_omerf$response,     lev)
cm_omerf_clm <- confusion_full(y_test_orig, pred_omerf_clm$response, lev)

cm_clm
cm_clmm
cm_ordforest
cm_omerf
cm_omerf_clm
```

# Feature importance — OMERF

```{r}
importance(omerf1$forest.model)
```

# Feature importance — iml

```{r}
if (requireNamespace("iml", quietly = TRUE)) {
  library(iml)
  library(ggplot2)

  mod_omerf <- Predictor$new(
    omerf1$forest.model,
    data = cov,
    y = omerf1$rf.target
  )
  imp_omerf <- FeatureImp$new(mod_omerf, loss = "mse")
  plot(imp_omerf)
}
```

# Partial dependence plots — iml

```{r}
mod11 <- Predictor$new(omerf1$forest.model, data = cov)
effs1 <- FeatureEffects$new(mod11, method = "pdp")
plot(effs1, nrows = 5, ncols = 4, fixed_y = FALSE)
```
