% A LaTeX template for ARTICLE version of the MSc Thesis submissions to 
% Politecnico di Milano (PoliMi) - School of Industrial and Information Engineering
%
% S. Bonetti, A. Gruttadauria, G. Mescolini, A. Zingaro
% e-mail: template-tesi-ingind@polimi.it
%
% Last Revision: October 2021
%
% Copyright 2021 Politecnico di Milano, Italy. Inc. NC-BY

\documentclass[11pt,a4paper]{article} 

%------------------------------------------------------------------------------
%	REQUIRED PACKAGES AND  CONFIGURATIONS
%------------------------------------------------------------------------------
% PACKAGES FOR TITLES
\usepackage{titlesec}
\usepackage{color}

% PACKAGES FOR LANGUAGE AND FONT
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc} % Font encoding
\usepackage{amsfonts}

% PACKAGES FOR IMAGES
\usepackage{graphicx}
\graphicspath{{Images/}}
\usepackage{eso-pic} % For the background picture on the title page
\usepackage{subfig} % Numbered and caption subfigures using \subfloat
\usepackage{caption} % Coloured captions
\usepackage{transparent}

% STANDARD MATH PACKAGES
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{bm}
\usepackage[overload]{empheq}  % For braced-style systems of equations

% PACKAGES FOR TABLES
\usepackage{tabularx}
\usepackage{longtable} % tables that can span several pages
\usepackage{colortbl}

% PACKAGES FOR ALGORITHMS (PSEUDO-CODE)
\usepackage{algorithm}
\usepackage{algorithmic}

% PACKAGES FOR REFERENCES & BIBLIOGRAPHY
\usepackage[colorlinks=true,linkcolor=black,anchorcolor=black,citecolor=black,filecolor=black,menucolor=black,runcolor=black,urlcolor=black]{hyperref} % Adds clickable links at references
\usepackage{cleveref}
\usepackage[square, numbers, sort&compress]{natbib} % Square brackets, citing references with numbers, citations sorted by appearance in the text and compressed
\bibliographystyle{plain} % You may use a different style adapted to your field

% PACKAGES FOR THE APPENDIX
\usepackage{appendix}

% PACKAGES FOR ITEMIZE & ENUMERATES 
\usepackage{enumitem}

% OTHER PACKAGES
\usepackage{amsthm,thmtools,xcolor} % Coloured "Theorem"
\usepackage{comment} % Comment part of code
\usepackage{fancyhdr} % Fancy headers and footers
\usepackage{lipsum} % Insert dummy text
\usepackage{tcolorbox} % Create coloured boxes (e.g. the one for the key-words)
\usepackage{listings}
\usepackage{hyperref}
\lstset{
  language=R,
  basicstyle=\ttfamily\small, % Tipo di carattere
  %commentstyle=\color{green}, % Colore dei commenti
  %keywordstyle=\color{blue}, % Colore delle parole chiave
  %stringstyle=\color{red}, % Colore delle stringhe
  showstringspaces=false, % Mostra gli spazi nelle stringhe
  breaklines=true, % Va a capo automaticamente se una linea è troppo lunga
  breakatwhitespace=true % Va a capo solo quando ci sono spazi vuoti
}

%-------------------------------------------------------------------------
%	NEW COMMANDS DEFINED
%-------------------------------------------------------------------------
% EXAMPLES OF NEW COMMANDS -> here you see how to define new commands
\newcommand{\bea}{\begin{eqnarray}} % Shortcut for equation arrays
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\e}[1]{\times 10^{#1}}  % Powers of 10 notation
\newcommand{\mathbbm}[1]{\text{\usefont{U}{bbm}{m}{n}#1}} % From mathbbm.sty
\newcommand{\pdev}[2]{\frac{\partial#1}{\partial#2}}
% NB: you can also override some existing commands with the keyword \renewcommand

%----------------------------------------------------------------------------
%	ADD YOUR PACKAGES (be careful of package interaction)
%----------------------------------------------------------------------------


%----------------------------------------------------------------------------
%	ADD YOUR DEFINITIONS AND COMMANDS (be careful of existing commands)
%----------------------------------------------------------------------------


% Do not change Configuration_files/config.tex file unless you really know what you are doing. 
% This file ends the configuration procedures (e.g. customizing commands, definition of new commands)
\input{Configuration_files/config}

% Insert here the info that will be displayed into your Title page 
% -> title of your work
\renewcommand{\title}{Ordinal Mixed-Effects Random Forest: an innovative statistical method to perform learning analytics.}
% -> author name and surname
\renewcommand{\author}{Giulia Bergonzoli}
% -> MSc course
\newcommand{\course}{Mathematical Engineering - Ingegneria Matematica}
% -> advisor name and surname
\newcommand{\advisor}{Dr. Chiara Masci}
% IF AND ONLY IF you need to modify the co-supervisors you also have to modify the file Configuration_files/title_page.tex (ONLY where it is marked)
\newcommand{\firstcoadvisor}{Dr.Ing. Lidia Rossi} % insert if any otherwise comment
%\newcommand{\secondcoadvisor}{Name Surname} % insert if any otherwise comment
% -> author ID
\newcommand{\ID}{994535}
% -> academic year
\newcommand{\YEAR}{2022-2023}
% -> abstract (only in English)
\renewcommand{\abstract}{In this work an innovative statistical method is proposed. The model is called Ordinal Mixed-Effect Random Forest (OMERF) and extends the use of random forest
to the analysis of hierarchical data to model ordinal categorical responses.
It preserves the flexibility and ability of modeling complex patterns of both categorical and continuous variables, typical of tree-based ensemble methods.
At the same time, OMERF takes into account the nested structure of hierarchical data, modeling the dependence structure that exists at the highest level of the hierarchy and allowing statistical inference on this structure.
In this study, a simulation is conducted to validate the performance of the proposed method and to compare OMERF against existing classical models.
The application of OMERF is exemplified in a case study focusing on modeling students at risk of failure from a prestigious high school in Milan, and generally on predicting their academic progress. 
The model is able to identify discriminating student characteristics and estimate the effect of each class to which students belong.}

% -> key-words (only in English)
\newcommand{\keywords}{ordinal response, hierarchical data, random forest, mixed-effects models, learning analytics}

%-------------------------------------------------------------------------
%	BEGIN OF YOUR DOCUMENT
%-------------------------------------------------------------------------
\begin{document}

%-----------------------------------------------------------------------------
% TITLE PAGE
%-----------------------------------------------------------------------------
% Do not change Configuration_files/TitlePage.tex (Modify it IF AND ONLY IF you need to add or delete the Co-advisors)
% This file creates the Title Page of the document
\input{Configuration_files/title_page}

\include{Thesis.tex}
\include{1.Introduction.tex}
\include{2.Literature_review.tex}
\include{3.Methods.tex}
\include{4.Simulation_study.tex}
\include{5.Case_study.tex}
\include{6.Conclusions.tex}
\include{7.Bibliography.tex}
\include{8.Appendix_A.tex}
%\include{9.Appendix_B.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%     ABSTRACT IN ITALIAN LANGUAGE AND ACKNOWLEDGMENTS     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\cleardoublepage

%-----------------------------------------------------------------------------
% SOMMARIO
%-----------------------------------------------------------------------------
\section*{Abstract in lingua italiana}
In questa tesi, proponiamo un innovativo metodo statistico. Il modello è chiamato Ordinal Mixed-Effect Random Forest (OMERF) e estende l'uso delle random forest all'analisi di dati gerarchici per modellare risposte categoriche ordinali.
Questo metodo preserva la flessibilità e la capacità di modellare strutture complesse sia per variabili categoriche che continue, tipiche dei metodi ensemble ad albero.
Allo stesso tempo, il modello OMERF tiene conto della struttura nidificata dei dati gerarchici, modellando la dipendenza che esiste al livello più alto della gerarchia e consentendo l'inferenza statistica su questa struttura.
In questo studio, è stata condotta una simulazione per validare le prestazioni del metodo proposto e confrontarlo con i più noti modelli già esistenti.
L'utilizzo dell'algoritmo OMERF è esemplificato dall'applicazione ad un caso reale incentrato sulla modellazione degli studenti a rischio di insuccesso di una prestigiosa scuola superiore di Milano e sulla previsione dei loro progressi accademici.
Il modello è in grado di identificare le caratteristiche discriminanti degli studenti e stimare l'effetto di ciascuna classe a cui gli studenti appartengono.
\vspace{15pt}
\begin{tcolorbox}[arc=0pt, boxrule=0pt, colback=bluePoli!60, width=\textwidth, colupper=white]
    \textbf{Parole chiave:} risposta ordinale, dati gerarchici, random forest, modelli a effetti misti, learning analytics 
\end{tcolorbox}

%-----------------------------------------------------------------------------
% ACKNOWLEDGEMENTS
%-----------------------------------------------------------------------------
\section*{Ringraziamenti}

Desidero esprimere la mia gratitudine alla relatrice di questa tesi, la professoressa Chiara Masci, per avermi offerto l'opportunità di contribuire a questo progetto di ricerca e per il suo continuo sostegno e incoraggiamento.
Un ringraziamento speciale è rivolto anche alla correlatrice Lidia Rossi, per la sua generosa disponibilità e per il suo costante supporto. \medskip

Ringrazio infinitamente i miei genitori e mio fratello, perchè mi hanno sempre motivato a dare il meglio, per il loro affetto e per la loro comprensione durante questo intenso periodo di studio. \medskip

Un grazie alle mie amiche più care, Chiara, Marzia, Rebecca e Beatrice, che hanno alleggerito i momenti più pesanti e mi sono sempre state vicine. \medskip

Infine, sono profondamente grata di aver incontrato tante persone che sono diventate amiche durante il mio percorso universitario. Insieme, abbiamo condiviso tanti ricordi e ci siamo supportati nei momenti più impegnativi. In particolare, grazie a Chiara e Michela per la vostra preziosa amicizia.


%-------------------------------------------------------------------------
%	END OF YOUR DOCUMENT
%-------------------------------------------------------------------------
\end{document}